---
title: "Mendelian randomisation: Critical Covid-19 as exposure for stroke (Additional analysis)"
author: "Verena Zuber"
date: "20/02/2021"
output: pdf_document
highlight: tango
---


# 1. Preparation


Load the MendelianRandomization package for the Mendelian randomization (MR) analysis, metafor for the forestplots, cowplot for plotting and knitr and markdown for compilation.
```{r echo = FALSE, message=FALSE, warning=FALSE}
library(MendelianRandomization)
library(metafor)
library(cowplot)
#library(knitr)
#library(markdown)
#library(rmarkdown)
```

Load the additional data file which is provided as Rdata file available from this github repository.
```{r message=FALSE, warning=FALSE}
load("additional.data.Rdata")
```

This data file contains three sets of data.
```{r message=FALSE, warning=FALSE}
#please uncomment to see full details
#str(additional.data$bidirectional.MR)
#str(additional.data$mvMR)
#str(additional.data$c19)
```


# 2. Bidirectional analysis: Considering ischemic stroke as exposure

Defining the mr_input objects which specify ischemic stroke and its subtypes as exposures for susceptibility to critical Covid-19.
```{r message=FALSE, warning=FALSE}
iv.ais = additional.data$bidirectional.MR$iv_A2.ais
iv.ces = additional.data$bidirectional.MR$iv_A2.ces
iv.las = additional.data$bidirectional.MR$iv_A2.las
iv.svs = additional.data$bidirectional.MR$iv_A2.svs
mr_ais = mr_input(by = iv.ais$beta_A2,byse = iv.ais$se_A2, 
 bx = iv.ais$beta_ais, bxse = iv.ais$se_ais)
mr_ces = mr_input(by = iv.ces$beta_A2,byse = iv.ces$se_A2, 
 bx = iv.ces$beta_ces, bxse = iv.ces$se_ces)
mr_las = mr_input(by = iv.las$beta_A2,byse = iv.las$se_A2, 
 bx = iv.las$beta_las, bxse = iv.las$se_las)
mr_svs = mr_input(by = iv.svs$beta_A2,byse = iv.svs$se_A2, 
 bx = iv.svs$beta_svs, bxse = iv.svs$se_svs)
```

Compute the inverse-variance weighted (IVW) MR estimate.
```{r message=FALSE, warning=FALSE}
ivw_ais = mr_ivw(mr_ais)
ivw_ces = mr_ivw(mr_ces)
ivw_las = mr_ivw(mr_las)
ivw_svs = mr_ivw(mr_svs)
```

IVW results including MR estimates, their standard error, confidence interval and heterogeneity statistics.
```{r echo = FALSE, message=FALSE, warning=FALSE}
tableIVW = matrix(nrow=4, ncol=7)
colnames(tableIVW) = c("Estimate","StdError","CILower","CIUpper","Pvalue","Q-stat", "Heter.Pvalue")
rownames(tableIVW) = c("Ischemic stroke","Cardioembolic stroke","Large artery stroke","Small vessel stroke")
tableIVW[1,] = c(ivw_ais$Estimate, ivw_ais$StdError, ivw_ais$CILower, ivw_ais$CIUpper, ivw_ais$Pvalue, ivw_ais$Heter.Stat)
tableIVW[2,] = c(ivw_ces$Estimate, ivw_ces$StdError, ivw_ces$CILower, ivw_ces$CIUpper, ivw_ces$Pvalue, ivw_ces$Heter.Stat)
tableIVW[3,] = c(ivw_las$Estimate, ivw_las$StdError, ivw_las$CILower, ivw_las$CIUpper, ivw_las$Pvalue, ivw_las$Heter.Stat)
tableIVW[4,] = c(ivw_svs$Estimate, ivw_svs$StdError, ivw_svs$CILower, ivw_svs$CIUpper, ivw_svs$Pvalue, ivw_svs$Heter.Stat)
tableIVW
```

In the manuscript we present the MR estimates and confidence intervalls on the odds ratio scale, where MR estimates represent the odds ratio for critical Covid-19 per unit increase in the log odds ratio of stroke phenotypes. 
```{r echo = FALSE, message=FALSE, warning=FALSE}
tableOR = as.data.frame(tableIVW)
tableOR$"Std Error" = NULL
tableOR$Estimate = exp(tableOR$Estimate)
tableOR$CILower = exp(tableOR$CILower)
tableOR$CIUpper = exp(tableOR$CIUpper)
tableOR
```

Figure 3: Forest plots of the bidirectional Mendelian randomization analysis illustrating the inverse-variance weighted Mendelian randomization estimate of ischemic stroke phenotypes (ischemic stroke, cardioembolic stroke, large artery stroke and small vessel stroke) as exposures on susceptibility to critical Covid-19.

```{r include = TRUE, fig.height = 6, fig.width = 8}
tableIVW = as.data.frame(tableIVW)
tableIVW$Pvalue = round(tableIVW$Pvalue, digits=3)
tableIVW$Heter.Pvalue = round(tableIVW$Heter.Pvalue, digits=3)
forest(x=tableIVW$Estimate,ci.lb=tableIVW$CILower, ci.ub=tableIVW$CIUpper,refline=1, 
 xlab="Odds ratio", slab=rownames(tableIVW),transf=exp, digits=3L, top=1, 
	ilab=cbind(tableIVW$Pvalue, tableIVW$Heter.Pvalue), ilab.xpos=c(0.615,0.705), 
	ilab.pos=4, xlim=c(0.35,1.48))
text(c(0.345,0.61,0.7,1.22), 4.4, pos=4, 
 c("Exposure","Pvalue","Het.Pvalue", "Estimate [CI]"),font=2)
```


# 3. Multivariable MR

Information on potential genetic confounder (smoking, obesity and inflammation) is contained in the following files.
```{r message=FALSE, warning=FALSE}
mvMR.ais = additional.data$mvMR$mvMR.ais
mvMR.ces = additional.data$mvMR$mvMR.ces
mvMR.las = additional.data$mvMR$mvMR.las
mvMR.svs = additional.data$mvMR$mvMR.svs
```

We performed multivariable MR to adjust for these potential genetic confounder, which we implemented in a multivariable weighted regression.
```{r message=FALSE, warning=FALSE}
#ais
lm.ais.uni = lm(beta_ais ~ beta_A2 -1, weights = se_ais^-2, data=mvMR.ais)
lm.ais.crp.uni = lm(beta_ais ~ beta_A2 -1, weights = se_ais^-2, 
 data = mvMR.ais[-which(is.na(mvMR.ais$beta_crp)==TRUE),])
lm.ais.smoking = lm(beta_ais ~ beta_A2 + beta_smoking -1, weights = se_ais^-2, 
 data=mvMR.ais)
lm.ais.bmi = lm(beta_ais ~ beta_A2 + beta_bmi -1, weights = se_ais^-2, data=mvMR.ais)
lm.ais.crp = lm(beta_ais ~ beta_A2 + beta_crp -1, weights = se_ais^-2, data=mvMR.ais)
lm.ais.joint = lm(beta_ais ~ beta_A2 + beta_smoking + beta_bmi + beta_crp -1, 
 weights = se_ais^-2, data=mvMR.ais)
#ces
lm.ces.uni = lm(beta_ces ~ beta_A2 -1, weights = se_ces^-2, data=mvMR.ces)
lm.ces.crp.uni = lm(beta_ces ~ beta_A2 -1, weights = se_ces^-2, 
 data = mvMR.ces[-which(is.na(mvMR.ces$beta_crp)==TRUE),])
lm.ces.smoking = lm(beta_ces ~ beta_A2 + beta_smoking -1, weights = se_ces^-2, 
 data=mvMR.ces)
lm.ces.bmi = lm(beta_ces ~ beta_A2 + beta_bmi -1, weights = se_ces^-2, data=mvMR.ces)
lm.ces.crp = lm(beta_ces ~ beta_A2 + beta_crp -1, weights = se_ces^-2, data=mvMR.ces)
lm.ces.joint = lm(beta_ces ~ beta_A2 + beta_smoking + beta_bmi + beta_crp -1, 
 weights = se_ces^-2, data=mvMR.ces)
#las
lm.las.uni = lm(beta_las ~ beta_A2 -1, weights = se_las^-2, data=mvMR.las)
lm.las.crp.uni = lm(beta_las ~ beta_A2 -1, weights = se_las^-2, 
 data = mvMR.las[-which(is.na(mvMR.las$beta_crp)==TRUE),])
lm.las.smoking = lm(beta_las ~ beta_A2 + beta_smoking -1, weights = se_las^-2, 
 data=mvMR.las)
lm.las.bmi = lm(beta_las ~ beta_A2 + beta_bmi -1, weights = se_las^-2, data=mvMR.las)
lm.las.crp = lm(beta_las ~ beta_A2 + beta_crp -1, weights = se_las^-2, data=mvMR.las)
lm.las.joint = lm(beta_las ~ beta_A2 + beta_smoking + beta_bmi + beta_crp -1, 
 weights = se_las^-2, data=mvMR.las)
#svs
lm.svs.uni = lm(beta_svs ~ beta_A2 -1, weights = se_svs^-2, data=mvMR.svs)
lm.svs.crp.uni = lm(beta_svs ~ beta_A2 -1, weights = se_svs^-2, 
 data = mvMR.svs[-which(is.na(mvMR.svs$beta_crp)==TRUE),])
lm.svs.smoking = lm(beta_svs ~ beta_A2 + beta_smoking -1, weights = se_svs^-2, 
 data=mvMR.svs)
lm.svs.bmi = lm(beta_svs ~ beta_A2 + beta_bmi -1, weights = se_svs^-2, data=mvMR.svs)
lm.svs.crp = lm(beta_svs ~ beta_A2 + beta_crp -1, weights = se_svs^-2, data=mvMR.svs)
lm.svs.joint = lm(beta_svs ~ beta_A2 + beta_smoking + beta_bmi + beta_crp -1, 
 weights = se_svs^-2, data=mvMR.svs)
```

Create the overview table for the multivariable MR analysis.
```{r message=FALSE, warning=FALSE}
mvMR.out = rbind(
summary(lm.ais.uni)$coefficients,summary(lm.ais.smoking)$coefficients[1,],
 summary(lm.ais.bmi)$coefficients[1,],summary(lm.ais.crp)$coefficients[1,],
 summary(lm.ais.joint)$coefficients[1,],
 #
 summary(lm.ces.uni)$coefficients,summary(lm.ces.smoking)$coefficients[1,], 
 summary(lm.ces.bmi)$coefficients[1,],summary(lm.ces.crp)$coefficients[1,],
 summary(lm.ces.joint)$coefficients[1,],
 #
 summary(lm.las.uni)$coefficients,summary(lm.las.smoking)$coefficients[1,], 
 summary(lm.las.bmi)$coefficients[1,],summary(lm.las.crp)$coefficients[1,],
 summary(lm.las.joint)$coefficients[1,],
 #
 summary(lm.svs.uni)$coefficients,summary(lm.svs.smoking)$coefficients[1,], 
 summary(lm.svs.bmi)$coefficients[1,],summary(lm.svs.crp)$coefficients[1,],
 summary(lm.svs.joint)$coefficients[1,]
)
rownames(mvMR.out) = c(
 "ais.total","ais.direct.adj.smoking","ais.direct.adj.bmi",
 "ais.direct.adj.crp","ais.direct.adj.all",
 "ces.total","ces.direct.adj.smoking","ces.direct.adj.bmi",
 "ces.direct.adj.crp","ces.direct.adj.all",
 "las.total","las.direct.adj.smoking","las.direct.adj.bmi",
 "las.direct.adj.crp","las.direct.adj.all",
 "svs.total","svs.direct.adj.smoking","svs.direct.adj.bmi",
 "svs.direct.adj.crp","svs.direct.adj.all"
)
#mvMR.out
mvMR.out = as.data.frame(mvMR.out)
mvMR.OR = as.data.frame(mvMR.out)
mvMR.OR$Estimate = exp(mvMR.out$Estimate)
mvMR.OR$CILower = exp(mvMR.out$Estimate-1.96*mvMR.out$"Std. Error")
mvMR.OR$CIUpper = exp(mvMR.out$Estimate+1.96*mvMR.out$"Std. Error") 
mvMR.OR$Pvalue = mvMR.OR$"Pr(>|t|)"
mvMR.OR$"Std. Error" = NULL
mvMR.OR$"t value" = NULL
mvMR.OR$"Pr(>|t|)" = NULL
mvMR.OR
```

Supplementary Figure 2: Forest plot contrasting the Mendelian randomization estimates and confidence intervals (CI) from univariable Mendelian randomization (total estimate) and multivariable Mendelian randomization accounting for potential pleiotropic pathways (direct estimate). 

```{r include = TRUE, fig.height = 12, fig.width = 8}
mvMR.out = mvMR.out[20:1,]
outcomes = c("Ischemic stroke","Cardioembolic stroke",
 "Large artery stroke","Small vessel stroke")
labels = rep(c("Total estimate", "Direct (adjusted for genetically predicted smoking)", 
 "Direct (adjusted for genetically predicted BMI)", 
 "Direct (adjusted for genetically predicted CRP)", 
 "Direct (adjusted for all)"),4)[20:1]
data.rma = rma(yi=mvMR.out$Estimate,sei=mvMR.out$"Std. Error") 
forest(data.rma,addfit=FALSE, rows=c(1:5, 7:11, 13:17, 19:23), ylim=c(0, 25), 
 slab=labels, atransf=exp,top=1, xlab="Odds ratio", digits=3L, xlim=c(-0.5,0.33))
 text(-0.51, c(24,18,12,6), pos=4, outcomes,font=2)
#points(rep(-0.495,4), c(24.04,18.04,12.04,6.04), pch=21,bg="black")
text(0.146, c(24), pos=4, c("Estimate and CI"),font=2)
```

In order to compare the model fit, we computed a likelihood ratio test which contrasts the residual sum of squares of the univariable with the multivariable MR model.
```{r message=FALSE, warning=FALSE}
#ais
lrt.ais.smoking = anova(lm.ais.uni, lm.ais.smoking)
lrt.ais.bmi = anova(lm.ais.uni, lm.ais.bmi)
lrt.ais.crp = anova(lm.ais.crp.uni, lm.ais.crp)
#ces
lrt.ces.smoking = anova(lm.ces.uni, lm.ces.smoking)
lrt.ces.bmi = anova(lm.ces.uni, lm.ces.bmi)
lrt.ces.crp = anova(lm.ces.crp.uni, lm.ces.crp)
#las
lrt.las.smoking = anova(lm.las.uni, lm.las.smoking)
lrt.las.bmi = anova(lm.las.uni, lm.las.bmi)
lrt.las.crp = anova(lm.las.crp.uni, lm.las.crp)
#svs
lrt.svs.smoking = anova(lm.svs.uni, lm.svs.smoking)
lrt.svs.bmi = anova(lm.svs.uni, lm.svs.bmi)
lrt.svs.crp = anova(lm.svs.crp.uni, lm.svs.crp)
```

Create the table to summarize the likelihood ratio test.
```{r message=FALSE, warning=FALSE}
lrt.test = rbind(
c(lrt.ais.smoking$RSS, lrt.ais.smoking$Df[2], lrt.ais.smoking$F[2], 
 lrt.ais.smoking$"Pr(>F)"[2]), 
c(lrt.ais.bmi$RSS, lrt.ais.bmi$Df[2], lrt.ais.bmi$F[2], 
 lrt.ais.bmi$"Pr(>F)"[2]),
c(lrt.ais.crp$RSS, lrt.ais.crp$Df[2], lrt.ais.crp$F[2], 
 lrt.ais.crp$"Pr(>F)"[2]),
c(lrt.ces.smoking$RSS, lrt.ces.smoking$Df[2], lrt.ces.smoking$F[2], 
 lrt.ces.smoking$"Pr(>F)"[2]), 
c(lrt.ces.bmi$RSS, lrt.ces.bmi$Df[2], lrt.ces.bmi$F[2], 
 lrt.ces.bmi$"Pr(>F)"[2]),
c(lrt.ces.crp$RSS, lrt.ces.crp$Df[2], lrt.ces.crp$F[2], 
 lrt.ces.crp$"Pr(>F)"[2]),
c(lrt.las.smoking$RSS, lrt.las.smoking$Df[2], lrt.las.smoking$F[2], 
 lrt.las.smoking$"Pr(>F)"[2]), 
c(lrt.las.bmi$RSS, lrt.las.bmi$Df[2], lrt.las.bmi$F[2], 
 lrt.las.bmi$"Pr(>F)"[2]),
c(lrt.las.crp$RSS, lrt.las.crp$Df[2], lrt.las.crp$F[2], 
 lrt.las.crp$"Pr(>F)"[2]),
c(lrt.svs.smoking$RSS, lrt.svs.smoking$Df[2], lrt.svs.smoking$F[2], 
 lrt.svs.smoking$"Pr(>F)"[2]), 
c(lrt.svs.bmi$RSS, lrt.svs.bmi$Df[2], lrt.svs.bmi$F[2], 
 lrt.svs.bmi$"Pr(>F)"[2]),
c(lrt.svs.crp$RSS, lrt.svs.crp$Df[2], lrt.svs.crp$F[2], 
 lrt.svs.crp$"Pr(>F)"[2])
)
colnames(lrt.test) = c("RSS1", "RSS2","DF","F","p-val")
rownames(lrt.test) = rep(c("smoking","bmi","crp"),4) 
lrt.test
```


# 4. Other Covid-19 definitions from the Covid-19 host genetics initiative

For the main analysis we considered the most severe definition of Covid-19 from the 5th release of the Covid-19 host genetics initiative https://www.covid19hg.org/results/r5/, which is defined as critical ill Covid-19, where a critical case is defined as an individual who was hospitalized with laboratory confirmed SARS-CoV-2 infection and required respiratory support or died. Here we present the following Covid-19 definitions: 
- B1: Hospitalized for Covid-19 versus controls with laboratory-confirmed Covid-19
- B2: Hospitalized for Covid-19 versus population controls
- C2: Reported Covid-19 infection versus population controls

Next we read in the data and define the mr_input objects for all three Covid-19 definitions.
```{r message=FALSE, warning=FALSE}
iv_B1 = additional.data$c19$iv_B1
iv_B2 = additional.data$c19$iv_B2
iv_C2 = additional.data$c19$iv_C2
#
mr_B1.ais = mr_input(bx = iv_B1$beta_B1,bxse = iv_B1$se_B1, 
 by = iv_B1$beta_ais, byse = iv_B1$se_ais)
mr_B1.ces = mr_input(bx = iv_B1$beta_B1,bxse = iv_B1$se_B1, 
 by = iv_B1$beta_ces, byse = iv_B1$se_ces)
mr_B1.las = mr_input(bx = iv_B1$beta_B1,bxse = iv_B1$se_B1, 
 by = iv_B1$beta_las, byse = iv_B1$se_las)
mr_B1.svs = mr_input(bx = iv_B1$beta_B1,bxse = iv_B1$se_B1, 
 by = iv_B1$beta_svs, byse = iv_B1$se_svs)
#
mr_B2.ais = mr_input(bx = iv_B2$beta_B2,bxse = iv_B2$se_B2, 
 by = iv_B2$beta_ais, byse = iv_B2$se_ais)
mr_B2.ces = mr_input(bx = iv_B2$beta_B2,bxse = iv_B2$se_B2, 
 by = iv_B2$beta_ces, byse = iv_B2$se_ces)
mr_B2.las = mr_input(bx = iv_B2$beta_B2,bxse = iv_B2$se_B2, 
 by = iv_B2$beta_las, byse = iv_B2$se_las)
mr_B2.svs = mr_input(bx = iv_B2$beta_B2,bxse = iv_B2$se_B2, 
 by = iv_B2$beta_svs, byse = iv_B2$se_svs)
#
mr_C2.ais = mr_input(bx = iv_C2$beta_C2,bxse = iv_C2$se_C2, 
 by = iv_C2$beta_ais, byse = iv_C2$se_ais)
mr_C2.ces = mr_input(bx = iv_C2$beta_C2,bxse = iv_C2$se_C2, 
 by = iv_C2$beta_ces, byse = iv_C2$se_ces)
mr_C2.las = mr_input(bx = iv_C2$beta_C2,bxse = iv_C2$se_C2, 
 by = iv_C2$beta_las, byse = iv_C2$se_las)
mr_C2.svs = mr_input(bx = iv_C2$beta_C2,bxse = iv_C2$se_C2, 
 by = iv_C2$beta_svs, byse = iv_C2$se_svs)
```

Perform inverse-variance weighted MR.
```{r message=FALSE, warning=FALSE}
ivw_B1.ais = mr_ivw(mr_B1.ais)
ivw_B1.ces = mr_ivw(mr_B1.ces)
ivw_B1.las = mr_ivw(mr_B1.las)
ivw_B1.svs = mr_ivw(mr_B1.svs)
#
ivw_B2.ais = mr_ivw(mr_B2.ais)
ivw_B2.ces = mr_ivw(mr_B2.ces)
ivw_B2.las = mr_ivw(mr_B2.las)
ivw_B2.svs = mr_ivw(mr_B2.svs)
#
ivw_C2.ais = mr_ivw(mr_C2.ais)
ivw_C2.ces = mr_ivw(mr_C2.ces)
ivw_C2.las = mr_ivw(mr_C2.las)
ivw_C2.svs = mr_ivw(mr_C2.svs)
```

Create the output table.
```{r message=FALSE, warning=FALSE}
tableC19 = matrix(nrow=12, ncol=7)
colnames(tableC19) = c("Estimate","StdError","CILower","CIUpper",
 "Pvalue","Q-stat", "Heter.Pvalue")
rownames(tableC19) = rep(c("Ischemic stroke","Cardioembolic stroke",
 "Large artery stroke","Small vessel stroke"),3)
#B1
tableC19[1,] = c(ivw_B1.ais$Estimate, ivw_B1.ais$StdError, 
 ivw_B1.ais$CILower, ivw_B1.ais$CIUpper, ivw_B1.ais$Pvalue, ivw_B1.ais$Heter.Stat)
tableC19[2,] = c(ivw_B1.ces$Estimate, ivw_B1.ces$StdError, 
 ivw_B1.ces$CILower, ivw_B1.ces$CIUpper, ivw_B1.ces$Pvalue, ivw_B1.ces$Heter.Stat)
tableC19[3,] = c(ivw_B1.las$Estimate, ivw_B1.las$StdError, 
 ivw_B1.las$CILower, ivw_B1.las$CIUpper, ivw_B1.las$Pvalue, ivw_B1.las$Heter.Stat)
tableC19[4,] = c(ivw_B1.svs$Estimate, ivw_B1.svs$StdError, 
 ivw_B1.svs$CILower, ivw_B1.svs$CIUpper, ivw_B1.svs$Pvalue, ivw_B1.svs$Heter.Stat)
#B2
tableC19[5,] = c(ivw_B2.ais$Estimate, ivw_B2.ais$StdError, 
 ivw_B2.ais$CILower, ivw_B2.ais$CIUpper, ivw_B2.ais$Pvalue, ivw_B2.ais$Heter.Stat)
tableC19[6,] = c(ivw_B2.ces$Estimate, ivw_B2.ces$StdError, 
 ivw_B2.ces$CILower, ivw_B2.ces$CIUpper, ivw_B2.ces$Pvalue, ivw_B2.ces$Heter.Stat)
tableC19[7,] = c(ivw_B2.las$Estimate, ivw_B2.las$StdError, 
 ivw_B2.las$CILower, ivw_B2.las$CIUpper, ivw_B2.las$Pvalue, ivw_B2.las$Heter.Stat)
tableC19[8,] = c(ivw_B2.svs$Estimate, ivw_B2.svs$StdError,  
 ivw_B2.svs$CILower, ivw_B2.svs$CIUpper, ivw_B2.svs$Pvalue, ivw_B2.svs$Heter.Stat)
#C2
tableC19[ 9,] = c(ivw_C2.ais$Estimate, ivw_C2.ais$StdError, 
 ivw_C2.ais$CILower, ivw_C2.ais$CIUpper, ivw_C2.ais$Pvalue, ivw_C2.ais$Heter.Stat)
tableC19[10,] = c(ivw_C2.ces$Estimate, ivw_C2.ces$StdError, 
 ivw_C2.ces$CILower, ivw_C2.ces$CIUpper, ivw_C2.ces$Pvalue, ivw_C2.ces$Heter.Stat)
tableC19[11,] = c(ivw_C2.las$Estimate, ivw_C2.las$StdError, 
 ivw_C2.las$CILower, ivw_C2.las$CIUpper, ivw_C2.las$Pvalue, ivw_C2.las$Heter.Stat)
tableC19[12,] = c(ivw_C2.svs$Estimate, ivw_C2.svs$StdError, 
 ivw_C2.svs$CILower, ivw_C2.svs$CIUpper, ivw_C2.svs$Pvalue, ivw_C2.svs$Heter.Stat)
#
#tableC19
```

In the manuscript we present the MR estimates and confidence intervalls on the odds ratio scale, where MR estimates represent the odds ratio for critical Covid-19 per unit increase in the log odds ratio of stroke phenotypes. 
```{r echo = FALSE, message=FALSE, warning=FALSE}
tableOR = as.data.frame(tableC19)
tableOR$"StdError" = NULL
tableOR$Estimate = exp(tableOR$Estimate)
tableOR$CILower = exp(tableOR$CILower)
tableOR$CIUpper = exp(tableOR$CIUpper)
tableOR
```

Supplementary Figure 4: Forest plot for other Covid-19 definitions
```{r include = TRUE, fig.height = 11, fig.width = 8}
data = tableC19[12:1,]
data = as.data.frame(data)
data$Pvalue = round(data$Pvalue, digits=3)
data$Heter.Pvalue = round(data$Heter.Pvalue, digits=3)
data.rma = rma(yi=as.numeric(data$Estimate),sei=as.numeric(data$StdError)) 
forest(data.rma,addfit=FALSE, rows=c(1:4, 6:9, 11:14), ylim=c(1, 15.5), 
 slab=rownames(tableC19), atransf=exp,top=1, xlab="Odds ratio", digits=3L, xlim=c(-0.45,0.9))
text(-0.46, c(15,10,5), pos=4, 
c("Hospitalized Covid-19 vs. not hospitalized Covid-19",
 "Hospitalized Covid-19 vs. population","Covid-19 vs. population"),font=2)
text(c(0.6), 15, pos=4, c("Estimate and CI"),font=2)
```

<!--
# for compilation use:
library(knitr)
library(markdown)
library(rmarkdown)
render("mr-additional-analysis.Rmd")
-->