---
title: "Mendelian randomisation: Critical Covid-19 as exposure for stroke"
author: "Verena Zuber"
date: "20/02/2021"
output: pdf_document
highlight: tango
---


# 1. Preparation


Load the MendelianRandomization package for the Mendelian randomization (MR) analysis, metafor for the forestplots, cowplot for plotting and knitr and markdown for compilation.
```{r echo = FALSE, message=FALSE, warning=FALSE}
library(MendelianRandomization)
library(metafor)
library(cowplot)
#library(knitr)
#library(markdown)
#library(rmarkdown)
```

Load the data which is provided in the Supplementary Table 2 available from this github repository.
```{r message=FALSE, warning=FALSE}
data = read.csv("SupplementaryTable2.csv", stringsAsFactors=FALSE)
```

This dataset contains summary-level information on genetic associations with the exposure:
- Critical Covid-19: beta_A2, se_A2, pval (from release 5 of https://www.covid19hg.org/results/r5/)
and ischemic stroke outcomes including
- Ischemic stroke (AIS): beta_ais, se_ais, p_ais
- Cardioembolic stroke (CES): beta_ces, se_ces, p_ces
- Large artery stroke (LAS): beta_las, se_las, p_las
- Small vessel stroke (SVS): beta_svs, se_svs, p_svs
```{r message=FALSE, warning=FALSE}
dim(data)
head(data, n=5)
```


# 2. Main analysis

Defining the mr_input objects for each exposure.
```{r message=FALSE, warning=FALSE}
mr_ais = mr_input(bx = data$beta_A2,bxse = data$se_A2, by = data$beta_ais, byse = data$se_ais)
mr_ces = mr_input(bx = data$beta_A2,bxse = data$se_A2, by = data$beta_ces, byse = data$se_ces)
mr_las = mr_input(bx = data$beta_A2,bxse = data$se_A2, by = data$beta_las, byse = data$se_las)
mr_svs = mr_input(bx = data$beta_A2,bxse = data$se_A2, by = data$beta_svs, byse = data$se_svs)
```

Compute the inverse-variance weighted (IVW) MR estimate.
```{r message=FALSE, warning=FALSE}
ivw_ais = mr_ivw(mr_ais)
ivw_ces = mr_ivw(mr_ces)
ivw_las = mr_ivw(mr_las)
ivw_svs = mr_ivw(mr_svs)
```

IVW results including MR estimates, their standard error, confidence interval and heterogeneity statistics.
```{r echo = FALSE, message=FALSE, warning=FALSE}
tableIVW = matrix(nrow=4, ncol=7)
colnames(tableIVW) = c("Estimate","StdError","CILower","CIUpper","Pvalue","Q-stat", "Heter.Pvalue")
rownames(tableIVW) = c("Ischemic stroke","Cardioembolic stroke","Large artery stroke","Small vessel stroke")
tableIVW[1,] = c(ivw_ais$Estimate, ivw_ais$StdError, ivw_ais$CILower, ivw_ais$CIUpper, ivw_ais$Pvalue, ivw_ais$Heter.Stat)
tableIVW[2,] = c(ivw_ces$Estimate, ivw_ces$StdError, ivw_ces$CILower, ivw_ces$CIUpper, ivw_ces$Pvalue, ivw_ces$Heter.Stat)
tableIVW[3,] = c(ivw_las$Estimate, ivw_las$StdError, ivw_las$CILower, ivw_las$CIUpper, ivw_las$Pvalue, ivw_las$Heter.Stat)
tableIVW[4,] = c(ivw_svs$Estimate, ivw_svs$StdError, ivw_svs$CILower, ivw_svs$CIUpper, ivw_svs$Pvalue, ivw_svs$Heter.Stat)
tableIVW
```

In the manuscript we present the MR estimates and confidence intervalls on the odds ratio scale, where MR estimates represent the odds ratio for critical Covid-19 per unit increase in the log odds ratio of stroke phenotypes. 
```{r echo = FALSE, message=FALSE, warning=FALSE}
tableOR = as.data.frame(tableIVW)
tableOR$"Std Error" = NULL
tableOR$Estimate = exp(tableOR$Estimate)
tableOR$CILower = exp(tableOR$CILower)
tableOR$CIUpper = exp(tableOR$CIUpper)
tableOR
```


Figure 2: Forest plot of MR estimates. 

```{r include = TRUE, fig.height = 6, fig.width = 8}
tableIVW = as.data.frame(tableIVW)
tableIVW$Pvalue = round(tableIVW$Pvalue, digits=3)
tableIVW$Heter.Pvalue = round(tableIVW$Heter.Pvalue, digits=3)
forest(x=tableIVW$Estimate,ci.lb=tableIVW$CILower, ci.ub=tableIVW$CIUpper,
	refline=1,xlab="Odds ratio",slab=rownames(tableIVW),transf=exp, digits=2L, top=1, 
	ilab=cbind(tableIVW$Pvalue, tableIVW$Heter.Pvalue), ilab.xpos=c(0.852,0.922), 
	ilab.pos=4, xlim=c(0.7,1.3))
text(c(0.7,0.85,0.92,1.166), 4.4, pos=4, 
 c("Outcome","Pvalue","Het.Pvalue", "Estimate [CI]"),font=2)
```

Supplementary Figure 1: Scatterplots of genetic association with the exposure (critical Covid-19) against the genetic association with the outcome (ischemic stroke).

```{r include = TRUE, fig.height = 14, fig.width = 14}
m1 = mr_plot(mr_ais, interactive = FALSE)
m2 = mr_plot(mr_ces, interactive = FALSE)
m3 = mr_plot(mr_las, interactive = FALSE)
m4 = mr_plot(mr_svs, interactive = FALSE)
plot_grid(m1,m2,m3,m4, labels = c('A', 'B', 'C', 'D'),nrow = 2, ncol = 2)
```


# 3. Sensitivity (Pleiotropy robust approaches) 

Compute as sensitivity pleiotropy-robust MR estimates.
```{r message=FALSE, warning=FALSE}
sensitivity_ais = mr_allmethods(mr_ais, method = "main")
sensitivity_ces = mr_allmethods(mr_ces, method = "main")
sensitivity_las = mr_allmethods(mr_las, method = "main")
sensitivity_svs = mr_allmethods(mr_svs, method = "main")
sensitivity = rbind(sensitivity_ais$Values,sensitivity_ces$Values, 
 sensitivity_las$Values, sensitivity_svs$Values)
colnames(sensitivity)[4:5] = c("CILower","CIUpper")
sensitivity
```

None of the MR-Egger estimates is significantly difference from zero.
```{r message=FALSE, warning=FALSE}
MREgger_intercept = sensitivity[c(5,10,15,20),]
rownames(MREgger_intercept) = c("Ischemic stroke","Cardioembolic stroke","Large artery stroke","Small vessel stroke")
MREgger_intercept
```

Transform estimates to the odds-ratio scale.
```{r echo = FALSE, message=FALSE, warning=FALSE}
sensitivityOR = list()
sensitivityOR$Method = sensitivity$Method
sensitivityOR$Estimate = exp(sensitivity$Estimate)
sensitivityOR$CILower = exp(sensitivity$CILower)
sensitivityOR$CIUpper = exp(sensitivity$CIUpper)
sensitivityOR$"P-value" = sensitivity$"P-value"
sensitivityOR = as.data.frame(sensitivityOR)
sensitivityOR = sensitivityOR[-c(5,10,15,20),]
sensitivityOR
```


# 4. Sensitivity (Genome-wide significant IVs)

Defining the mr_input objects for each exposure on genome-wide significant genetic variants as instrumental variables only.
```{r message=FALSE, warning=FALSE}
data = data[data$pval<5e-8,]
dim(data)
mr2_ais = mr_input(bx = data$beta_A2,bxse = data$se_A2, by = data$beta_ais, byse = data$se_ais)
mr2_ces = mr_input(bx = data$beta_A2,bxse = data$se_A2, by = data$beta_ces, byse = data$se_ces)
mr2_las = mr_input(bx = data$beta_A2,bxse = data$se_A2, by = data$beta_las, byse = data$se_las)
mr2_svs = mr_input(bx = data$beta_A2,bxse = data$se_A2, by = data$beta_svs, byse = data$se_svs)
ivw2_ais = mr_ivw(mr2_ais)
ivw2_ces = mr_ivw(mr2_ces)
ivw2_las = mr_ivw(mr2_las)
ivw2_svs = mr_ivw(mr2_svs)
```

IVW results including MR estimates, their standard error, confidence interval and heterogeneity statistics.
```{r echo = FALSE, message=FALSE, warning=FALSE}
table2IVW = matrix(nrow=4, ncol=7)
colnames(table2IVW) = c("Estimate","StdError","CILower","CIUpper","Pvalue","Q-stat", "Heter.Pvalue")
rownames(table2IVW) = c("Ischemic stroke","Cardioembolic stroke","Large artery stroke","Small vessel stroke")
table2IVW[1,] = c(ivw2_ais$Estimate, ivw2_ais$StdError, ivw2_ais$CILower, ivw2_ais$CIUpper, ivw2_ais$Pvalue, ivw2_ais$Heter.Stat)
table2IVW[2,] = c(ivw2_ces$Estimate, ivw2_ces$StdError, ivw2_ces$CILower, ivw2_ces$CIUpper, ivw2_ces$Pvalue, ivw2_ces$Heter.Stat)
table2IVW[3,] = c(ivw2_las$Estimate, ivw2_las$StdError, ivw2_las$CILower, ivw2_las$CIUpper, ivw2_las$Pvalue, ivw2_las$Heter.Stat)
table2IVW[4,] = c(ivw2_svs$Estimate, ivw2_svs$StdError, ivw2_svs$CILower, ivw2_svs$CIUpper, ivw2_svs$Pvalue, ivw2_svs$Heter.Stat)
table2IVW
```

In the manuscript we present the MR estimates and confidence intervalls on the odds ratio scale, where MR estimates represent the odds ratio for critical Covid-19 per unit increase in the log odds ratio of stroke phenotypes. 
```{r echo = FALSE, message=FALSE, warning=FALSE}
table2OR = as.data.frame(table2IVW)
table2OR$"Std Error" = NULL
table2OR$Estimate = exp(table2OR$Estimate)
table2OR$CILower = exp(table2OR$CILower)
table2OR$CIUpper = exp(table2OR$CIUpper)
table2OR
```



Supplementary Figure 3: Forest plot of MR estimates (genome-wide significant genetic variants as instrumental variables).

```{r include = TRUE, fig.height = 6, fig.width = 8}
table2IVW = as.data.frame(table2IVW)
table2IVW$Pvalue = round(table2IVW$Pvalue, digits=3)
table2IVW$Heter.Pvalue = round(table2IVW$Heter.Pvalue, digits=3)
forest(x=table2IVW$Estimate,ci.lb=table2IVW$CILower, ci.ub=table2IVW$CIUpper,
	refline=1,xlab="Odds ratio",slab=rownames(table2IVW),transf=exp, digits=3L, top=1, 
	ilab=cbind(table2IVW$Pvalue, table2IVW$Heter.Pvalue), ilab.xpos=c(0.835,0.89), 
	ilab.pos=4, xlim=c(0.7,1.28))
text(c(0.7,0.831,0.886,1.148), 4.4, pos=4, 
 c("Outcome","Pvalue","Het.Pvalue", "Estimate [CI]"),font=2)
```



<!--
# for compilation use:
library(knitr)
library(markdown)
library(rmarkdown)
render("mr-covid19-stroke.Rmd")
-->